Stateful Delta Lenses

### 2018-02-15 papers/SDLenses/main.tex
Made changes

\usepackage{amsmath, amssymb, amsthm, enumerate, graphicx, centernot, tikz,
array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd, listings}

@@

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\get}{\ensuremath{\kw{get}}}
\newcommand{\pput}{\ensuremath{\kw{put}}}
\newcommand{\create}{\ensuremath{\kw{create}}}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}}
\newcommand{\const}{\ensuremath{\kw{const}}}
\newcommand{\default}{\ensuremath{\kw{default}}}
\newcommand{\pos}{\ensuremath{\kw{posit}}}
\newcommand{\length}{\ensuremath{\kw{length}}}
\newcommand{\match}{\ensuremath{\kw{match}}}

@@

This idea of lenses carrying a notion of updates and update propagation has been
explored in various ways. First, various categorical approaches have been
explored under the name of {\em delta lenses}
\cite{diskin2011state,johnson2017symmetric,pacheco2012delta,johnson2012lenses}
Delta lenses are lenses defined in an appropriate category, where sources and
views are the objects of the category, and the arrows are deltas (or updates)
between the objects of the category. In this setting, lens transformations are
functors which map objects to objects as well as deltas to deltas. Categorical
delta lenses are thus similar to the SD-lenses described in this paper in that
they map states to states as well as updates to updates. The main difference
between SD-lenses and categorical delta lenses is that SD-lenses are
set-based rather than categorical and arguably simpler to understand. Moreover,
only \cite{pacheco2012delta} defines a fairly rich set of combinators for delta
lenses over inductive types in the same vane as SD-lenses.

>>>>

\usepackage{amsmath, amssymb, amsthm, enumerate, graphicx, centernot, tikz,
array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd, listings,
xspace}

@@

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}\xspace}
\newcommand{\get}{\ensuremath{\kw{get}}\xspace}
\newcommand{\pput}{\ensuremath{\kw{put}}\xspace}
\newcommand{\create}{\ensuremath{\kw{create}}\xspace}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}\xspace}
\newcommand{\const}{\ensuremath{\kw{const}}\xspace}
\newcommand{\default}{\ensuremath{\kw{default}}\xspace}
\newcommand{\pos}{\ensuremath{\kw{posit}}\xspace}
\newcommand{\length}{\ensuremath{\kw{length}}\xspace}
\newcommand{\match}{\ensuremath{\kw{match}}\xspace}

@@

This idea of lenses carrying a notion of updates and update propagation has been
explored in various ways. First, various categorical approaches have been
explored under the name of {\em delta lenses}
\cite{diskin2011state,diskin2011asymmetric,johnson2017symmetric,pacheco2012delta,johnson2012lenses} Delta lenses are lenses defined in an appropriate category, where sources and
views are the objects of the category, and the arrows are deltas (or updates)
between the objects of the category. In this setting, lens transformations are
functors which map objects to objects as well as deltas to deltas. Categorical
delta lenses are thus similar to the SD-lenses described in this paper in that
they map states to states as well as updates to updates.

Our SD-lenses are set-based version of the delta lenses introduced by Pacheco et
al in \cite{pacheco2012delta}. Pacheco et al. define a delta lens $\ell$ (d-lens
for short), denote by $\ell: SA \trianglerighteq_{\blacktriangle} VB$ to be a
bidirectional transformations that comprises four total functions:
\begin{align*}
\get &: SA \longrightarrow SB\\
\pput &: \forall(v, s) : VB \times SA. (v \triangle \get \;s) \longrightarrow
SA\\
\get_{\blacktriangle} &: \forall\{s':SA, s:SA\}. (s' \triangle s)
\longrightarrow (\get \; s \triangle \get \; s)\\
\pput_{\blacktriangle} &:\forall\{(v, s):VB \times SA\},d:(v \triangle \get
\;s).(\pput \; (v, s) \; d) \triangle s
\end{align*}
The d-lens is called {\em well-behaved} iff it satisfies the following
propertes:
\begin{align*}
\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
\tag{$\text{PUTGET}_{\blacktriangle}$}\\
\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
\end{align*}
In the above definition, the state-based component of the d-lens is given by
the functions get, that computes a view of a source value, and put, that takes
a pair containing a modified view and an original source, together with a delta
from the modified view to the original view, and returns a new modified value.
The delta-based function $\get_{\blacktriangle}$ translates a source delta into
a delta between views produced by $\get$, and $\pput_{\blacktriangle}$ receives
a view delta and computes a delta from the new source produced by put to the
original source. Properties PUTGET and GETPUT are the traditional state-based
ones: view-to-view roundtrips preserve view modifications; and $\pput$ must
preserve the original source for identity updates.
$\text{PUTGET}_{\blacktriangle}$ and $\text{PUTID}_{\blacktriangle}$ denote
similar laws on deltas: view-to-view roundtrips preserve view updates; and
$\text{PUTID}_{\blacktriangle}$ must preserve identity updates.

SD-lenses are essentially equivalent to well-behaved d-lenses in the sense of
Pacheco et al.
by mapping the state-based and delta based components of an SD-lens to
their respective state and delta based components in a well-behaved d-lens.
The axiom $\text{PUTID}_{\blacktriangle}$ combined with the dependent type of
$\pput_{\blacktriangle}$ hence implies the axiom
$$(\partial (S, V)).\get \; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
\ell.\get \; s)$$ 
for SD-lenses. Pacheco et al. also observe that their formulation is equivalent
to the well-behaved d-lenses defined by Diskin et al. in
\cite{diskin2011asymmetric}, hence our SD-lenses are essentially equivalent to
these lenses as well.
